Experience Report: Distributed Computing for the Common Man

### 2010-03-29 doc/icfp-2010/main.tex
factory -> functory

\newcommand{\Ocaml}{OCaml}
\newcommand{\JoCaml}{Jo{\&\!}Caml}
\newcommand{\unix}{\textsc{Unix}}

@@

The application domain of such a distributed computing library is
manyfold.  The most striking application, as also relevant to our
research endeavours, is in validating thousands of verification
conditions using automated theorem provers. Having noted this, it
should be apparent that the library finds use in efficient execution
of any large-scale expensive computation that is parallelizable.
% what it is and what it is not
% \begin{itemize}
% \item it is not a library which does parallelization
% \item it is a library for reliable distributed execution of
%   parallelizable computation 
% \end{itemize}
% inspired by Google's MapReduce~\cite{mapreduce} (itself inspired by functional
% programming, ironically) ;  however, there are differences:
% \begin{itemize}
% \item polymorphic, higher-order API
% \item more generic (cores/network)
% \item does not focus on association lists
% \item we don't have data locality (future work)
% \end{itemize}
% what is the target audience/applications: 
% for instance, use if automatic provers on thousands of verification
% conditions, on computing infrastructures which can be
% \begin{itemize}
% \item a single machine with multiple cores, possibly remote
% \item several machines, small or large in number, over a network
% \end{itemize}
\paragraph{Distributed Computing.}
A typical distributed computing library provides the following (we
borrow some terminology from Google's Mapreduce~\cite{mapreduce}):
\begin{itemize}
\item A notion of a \emph{task} which denotes the atomic computation to be
  performed in a distributed manner;
\item A set of machines or programs called \emph{workers} that perform
  the tasks, producing results;
\item A single program called a \emph{master} that is in charge
  of distributing the tasks among the workers and managing results
  produced by the workers.
\end{itemize}
In addition to the above, distributed computing environments also
implement mechanisms for fault-tolerance, efficient storage and
distribution of tasks. This is required to handle network failures
that may occur, as well as to optimize the usage of machines in the
network. Another concern of importance is the transmission of messages
over the network. This requires efficient serialization (or
marshalling) of data that can be reliably reconstructed when
transmitted over different computing environments.  It is desirable to
maintain architecture independence while transmitting marshalled data,
as machines in a distributed computing environment often run on
different hardware architectures and make use of varying software
platforms. For example, endianness may be different across machines on
the network.

@@

In this section, we show the potential of our library using two 
computationally-extensive examples.

@@

Here we demonstrate the potential of our library for our application
needs as mentioned in the introduction. Our computing
infrastructure consists of 3 machines with 4, 8 and 8 cores
respectively. The benchmark we use for
experimentation consists of 80 challenging verification conditions
(VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
VC is stored in a file, which is accessible over
NFS. The purpose of the experiment is to check the validity of each VC
using several automated provers (namely Alt-Ergo, Simplify and Z3),
which are all installed on each of the machines in use.

The master program proceeds by reading the 80 file names, turning them
into tasks by multiplying them by the number of provers (240 tasks in total).
Each worker in turn invokes the given prover on the given file.
We set a timeout limit of 10 minutes.
Each task completes with one of the four possible outcomes: \emph{valid},
\emph{cannot-decide} (depending on
whether the VC is valid or undecided by the prover), 
\emph{timeout} and \emph{failure}.
The figure below shows the total time in minutes spent by each prover
for each possible outcomes.
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    prover   & valid & cannot & timeout & failure
    \\\hline\hline
    Z3       &   4.20 &   0.00 &    160  &  20.0    
    \\\hline
    Alt-ergo &   3.60 &   0.57 &    190  &   8.2     
    \\\hline
    Simplify &   0.02 &   0.00 &     40  &   1.0   
    \\\hline
  \end{tabular}
\end{center}
These figures sum up to more than 7 hours if provers were executed
sequentially. However, using our library and our 3-machine
infrastructure, it completes in 1 hour and 5 minutes, giving us a
speedup of 6.6$\times$.

% timeout = 600
%            |  valid       | invalid      | cannot       | timeout      | failure
% ---------------------------------------------------------------------------------------
%         z3 |  42 (251.25s) |   0 ( 0.00s) |   0 ( 0.00s) |  16 (9548.79s) |  22 (1223.81s) | 
%   alt-ergo |  40 (217.55s) |   0 ( 0.00s) |  19 (34.29s) |  19 (11376.70s) |   2 (489.05s) | 
%   simplify |  32 ( 0.95s) |   0 ( 0.00s) |  34 ( 0.25s) |   4 (2394.60s) |  10 (62.76s) | 
% 0.43user 0.33system 1:04:57elapsed 0%CPU (0avgtext+0avgdata 0maxresident)k

% total amount of time = 25630 seconds = 7 hours 7 minutes 10 seconds
% ratio = 6.57

% timeout = 120
%            |  valid       | invalid      | cannot       | timeout      | failure
% ---------------------------------------------------------------------------------------
%         z3 |  41 (211.08s) |   0 ( 0.00s) |   0 ( 0.00s) |  20 (2383.89s) |  19 (262.30s) | 
%   alt-ergo |  40 (237.00s) |   0 ( 0.00s) |  19 (34.45s) |  21 (2514.42s) |   0 ( 0.00s) | 
%   simplify |  32 ( 0.98s) |   0 ( 0.00s) |  34 ( 0.31s) |   5 (598.92s) |   9 (18.98s) |

@@

Closest to the approach in this paper is Yohann Padioleau's mapreduce
implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
top of OcamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
protocol for message passing.  On one hand, we have less flexibility
w.r.t. deployment of the user program; on the other hand, we provide a
more generic API together with fault tolerance.  There are other
distributed computing libraries on top of which one could implement
the library discussed in this paper. \JoCaml~\cite{jocaml} is one of
them. However, \JoCaml\ does not allow arbitrary closures to be passed
between machines, which would limit the genericity of the
API. Additionally, \JoCaml\ does not provide fault tolerance, which is
indispensible in a distributed setting.
% TODO check with Louis

@@

\nocite{*}
\bibliographystyle{abbrvnat}
\bibliography{./biblio}

>>>>

\newcommand{\Ocaml}{OCaml}
\newcommand{\functory}{\textsf{Functory}}
\newcommand{\JoCaml}{Jo{\&\!}Caml}
\newcommand{\unix}{\textsc{Unix}}

@@

The application domain of such a distributed computing library is
manyfold.  The most striking application, as also relevant to our
research endeavours, is in validating thousands of verification
conditions using automated theorem provers. Having noted this, it
should be apparent that the library finds use in efficient execution
of any large-scale expensive computation that is parallelizable.
% what it is and what it is not
% \begin{itemize}
% \item it is not a library which does parallelization
% \item it is a library for reliable distributed execution of
%   parallelizable computation 
% \end{itemize}
% inspired by Google's MapReduce~\cite{mapreduce} (itself inspired by functional
% programming, ironically) ;  however, there are differences:
% \begin{itemize}
% \item polymorphic, higher-order API
% \item more generic (cores/network)
% \item does not focus on association lists
% \item we don't have data locality (future work)
% \end{itemize}
% what is the target audience/applications: 
% for instance, use if automatic provers on thousands of verification
% conditions, on computing infrastructures which can be
% \begin{itemize}
% \item a single machine with multiple cores, possibly remote
% \item several machines, small or large in number, over a network
% \end{itemize}
\paragraph{Distributed Computing.}
A typical distributed computing library provides the following (we
borrow some terminology from Google's Mapreduce~\cite{mapreduce}):
\begin{itemize}
\item A notion of a \emph{task} which denotes the atomic computation to be
  performed in a distributed manner;
\item A set of machines or programs called \emph{workers} that perform
  the tasks, producing results;
\item A single program called a \emph{master} that is in charge
  of distributing the tasks among the workers and managing results
  produced by the workers.
\end{itemize}
In addition to the above, distributed computing environments also
implement mechanisms for fault-tolerance, efficient storage and
distribution of tasks. This is required to handle network failures
that may occur, as well as to optimize the usage of machines in the
network. Another concern of importance is the transmission of messages
over the network. This requires efficient 
\emph{marshalling} of data, that is encoding and decoding of data 
for transmission over different computing environments.  It is desirable to
maintain architecture independence while transmitting marshalled data,
as machines in a distributed computing environment often run on
different hardware architectures and make use of varying software
platforms. For example, endianness may be different across machines on
the network.

@@

Our library, \functory, 
In this section, we show the potential of our library using two 
computationally-extensive examples.

@@

Here we demonstrate the potential of our library for our application
needs as mentioned in the introduction. Our computing
infrastructure consists of 3 machines with 4, 8 and 8 cores
respectively. The benchmark we use for
experimentation consists of around 80 challenging verification conditions
(VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
VC is stored in a file, which is accessible over
NFS. The purpose of the experiment is to check the validity of each VC
using several automated provers (namely Alt-Ergo, Simplify, Z3 and CVC3),
which are all installed on each of the machines in use.

The master program proceeds by reading the file names, turning them
into tasks by multiplying them by the number of provers (more than
300 tasks in total).
Each worker in turn invokes the given prover on the given file.
We set a timeout limit of 1 minute.
Each task completes with one of the four possible outcomes: \emph{valid},
\emph{cannot-decide} (depending on
whether the VC is valid or undecided by the prover), 
\emph{timeout} and \emph{failure}.
The figure below shows the total time in minutes spent by each prover
for each possible outcomes.
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    prover   & valid & cannot & timeout & failure
    \\\hline\hline
    Alt-ergo &  24.7  &  2.4  & 1260.0 &   0.0       
    \\\hline
    Simplify &   0.6  &  0.3  &  300.0 &  43.0   
    \\\hline
    Z3       &  98.6  &  0.0  & 1200.0 &  66.3    
    \\\hline
    CVC3     &  46.7  & 28.0  &  720.0 &   0.0   
    \\\hline
  \end{tabular}
\end{center}
These figures sum up to more than 1 hour if provers were executed
sequentially. However, using our library and our 3-machine
infrastructure, it completes in 4 minutes and 8 seconds, giving us a
speedup of more than 15$\times$.

@@

Closest to the approach in this paper is Yohann Padioleau's mapreduce
implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
top of OcamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
protocol for message passing.  On one hand, we have less flexibility
w.r.t. deployment of the user program; on the other hand, we provide a
more generic API together with fault tolerance.  There are other
distributed computing libraries on top of which one could implement
the library discussed in this paper. \JoCaml~\cite{jocaml} is one of
them. However, \JoCaml\ does not provide fault tolerance, which is
indispensible in a distributed setting. The user has to include code
for fault tolerance, as already demonstrated in some \JoCaml\
experiments~\cite{mandel2008}.

@@

\bibliographystyle{abbrvnat}
\bibliography{./biblio}
